---
title: 'N741: Linear and Logistic Regression'
author: "Melinda K. Higgins, PhD."
date: "February 8, 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

## Initial R Chunk - document and environment setup

```{r setup, include=FALSE}
# initial R chunk - set global knitr options
# for now echo=TRUE so ALL R chunks will be shown
# in your final output.
#
# When you are ready to do a final document NOT showing
# the R code - run this with echo = FALSE instead.
# TRY SETTING echo = FALSE and see how the report changes.
knitr::opts_chunk$set(echo = TRUE)

# In this inital R chunk, this is the best place to go ahead
# and load ALL of the packages your R markdown document
# will need. Then you can then not have to worry about adding 
# the library() commands later on.
# Tis also helps anyone reading your code to know
# right at the beginning which packages are needed to
# execute your code and compile your document. Thus, they
# can go ahead and install the packages needed (which is done
# outside of and BEFORE the RMD file is compiled)

# packages needed for this RMD file
library(tidyverse)
```

## Setup `CMC` dataset from UCI Data Repository

We've covered this before, but the following code will read in the modified dataset and apply the factor level coding and labels.

```{r}
# The modified dataset is comma delimited but does have
# slightly different column names from what we had before.
cmc <- read_csv("cmcmodfinal.csv")
```

### Apply the codebook - variable names and coding used

Apply variable names to the 10 columns of data in `cmc`.

```{r}
# since we want to reuse the code we had before, let's
# assign the variables names we had before to the 10 columns
names(cmc) <- c("WifeAge", "WifeEd", "HusbEd", "NumChild",
                "WifeRel", "WifeWork", "HusbOcc", "SOLindex", 
                "Media", "Contraceptive")
```

The next code chunk is to add the labels for "factor" levels for some of the variables (i.e. we are creating factors).

Use the code below to update all of the variables.

```{r}
# update Wife Education as a factor, assign the 
# levels and the labels for each level
cmc$WifeEd <- factor(cmc$WifeEd,
                     levels = c(1,2,3,4),
                     labels = c("low","med low","med high","high"))

# do the remaining variables
cmc$HusbEd <- factor(cmc$HusbEd,
                     levels = c(1,2,3,4),
                     labels = c("low","med low","med high","high"))

cmc$WifeRel <- factor(cmc$WifeRel,
                      levels = c(0,1),
                      labels = c("Non-Islam","Islam"))

# Note: The documentation does state that
# 0=yes and 1=no which seems incorrect...
cmc$WifeWork <- factor(cmc$WifeWork,
                       levels = c(0,1),
                       labels = c("Yes","No"))

cmc$HusbOcc <- factor(cmc$HusbOcc,
                      levels = c(1,2,3,4),
                      labels = c("1","2","3","4"))

cmc$SOLindex <- factor(cmc$SOLindex,
                       levels = c(1,2,3,4),
                       labels = c("low","med low","med high","high"))

cmc$Media <- factor(cmc$Media,
                    levels = c(0,1),
                    labels = c("Good","Not Good"))

cmc$Contraceptive <- factor(cmc$Contraceptive,
                            levels = c(1,2,3),
                            labels = c("No-use","Long-term","Short-term"))
```

### Save a copy of the `cmc` dataset

The `save()` command will save a copy of your dataset (the `cmc` object) as an `.RData` file which is easily read by R. The associated command to then read the data back in is the `load()` command. This will be helpful shortly when we run the demo in "R Commander" (`Rcmdr` package).

```{r}
# save the cmc dataset with the updated variable names
# and associated factor levels and labeling applied.
save(cmc, file="cmc.RData")
```

Do a quick `summary()` of the data - what do you notice?

```{r}
summary(cmc)
```

Issues noted:

* The minimum value for Wife's Age is now -9 AND the max is 114 (outliers?)
* Husband's education now lists 461 NA's (missing values)
* There are also 461 missing values (NAs) for SOLindex and Media.

NOTE: If you look at the dataset, you'll notice that the 461 subjects are the same ones missing Husband's education, SOL index and Media - but this is usually not the case - missing data is often intermittent and will vary across variables.

```{r}
hist(cmc$WifeAge)
sum(cmc$WifeAge < 18)
sum(cmc$WifeAge > 90)
hist(cmc$WifeAge[cmc$WifeAge<20])

cmc %>%
  filter(WifeAge < 20) %>%
  select(WifeAge) %>%
  ggplot(aes(WifeAge)) +
    geom_histogram(binwidth=1)

cmc %>%
  filter(WifeAge > 40) %>%
  select(WifeAge) %>%
  ggplot(aes(WifeAge)) +
    geom_histogram(binwidth=1)
  
# add a column that removes these outliers
# use the mutate() function, set ages < 16 and > 60 to missing NA
cmc <- cmc %>% 
  mutate(WifeAgeFix = ifelse(WifeAge < 16, NA,
                             ifelse(WifeAge > 60, NA, WifeAge)))

hist(cmc$WifeAgeFix)
summary(cmc$WifeAgeFix)

cmc %>%
  ggplot(aes(WifeAgeFix)) +
    geom_histogram(binwidth=1)

summary(cmc$HusbEd)

cmc %>%
  ggplot(aes(HusbEd)) +
    geom_bar()

cmc %>%
  filter(!is.na(HusbEd)) %>%
  ggplot(aes(HusbEd)) +
    geom_bar()


save(cmc, file="cmc.RData")

library(Rcmdr)
```


## wine data


